The language of generalization: Probability, vagueness, and interaction

### 2018-10-10 paper/comgen-paper.Rmd
polish worked examples

Model predictions depend upon the uncertain threshold model's interpretations, which are highly sensitive to background knowledge.
Background knowledge may be richly structured, reflecting intuitive theories about the underlying causes of different kinds of properties [cf., @Leslie2007].
This model posits, however, that the only impact of structured knowledge on truth judgments is in their implications for beliefs about prevalence, formalized in terms of a prevalence prior $P(p)$. 
\ndg{i’d save this for expt sections, and here just say that we posit intuitive priors for illustrations and introduce a method to actually elicit them below}
\mht{without this, we don't have an explanation for what the prevalence prior is, which is important for understanding the examples... i removed stuff about Betas, but check this again...}
The prevalence prior can be thought of a distribution over the prevalence of a feature among alternative categories (Figure\ \@ref(fig:simulations)).
<!-- For example, think about five of your favorite kinds of animal: What percentage of each of them *are female*? What percentage of each of them *lay eggs*? -->
Figure\ \@ref(fig:simulations) (row 1) shows intuitive prevalence priors for six different properties (we introduce methods for empirically eliciting these priors in the experimental section).
The *lay eggs* distribution is bimodal, with substantial probability mass near 0 (since many animal categories do not have egg-layers) and a secondary component peaked around 0.5, because among the animals with egg-layers, only the female members of the category have the property.
Conversely, the distribution over the prevalence of *being female* is unimodal and centered at 0.5, because almost all animals have female members in that proportion.

@@

```{r workedExamples, results="asis"}
worked.examples <- data.frame(
  Example = c("1. Dogs bark",
              "2. Kangaroos have spots",
              "3. Sharks don't eat people",
              "4. Robins lay eggs",
              "5. Robins are female",
              "6. Mosquitos carry malaria", 
              "7. Mary handles the mail from Antarcita",
              "8. Supreme Court Justices have even social security numbers",
              "9. Elephants live in Africa and Asia"),
  `referent prevalence` = c("95%", "5%", "85%", "50%", "50%", "5%", "0 times", "100%", "50%/50%"),
  `intuitive truth value` = c(
    "T", "F", "F", "T", "U",
    "T", "T", "F", "T" 
  ),
  `potential issue` = c(
    "Not all dogs bark",
    "Some kangaroos could have spots",
    "Most sharks don't eat people",
    "True, even though only female robins (50%) lay eggs",
    "Not true, but same number that lay eggs (50%)",
    "Very few mosquitos actually carry malaria",
    "True even if there has never been any mail from Antarcita",
    "False even if 100% have even social security numbers",
    "Impossible for most to live in Africa and most to live in Asia"
  )
) %>%
  rename("Referent prevalence" = referent.prevalence,
         "Intuitive truth value" = intuitive.truth.value,
         "Issue" = potential.issue)

@@

Often the first observation with generics is that they appear to behave like universal quantifiers (e.g., "All") that permit exceptions.
That is, not all dogs bark, but still "Dogs bark" is true.
The uncertain threshold account immediately accomodates this example because there is no single, fixed threshold beyond which the generalization becomes true.
Rather, listeners have uncertainty about the threshold *a priori* which leads to graded interpretations (e.g., "Dogs bark" meaning *almost all* dogs bark; Figure\ \@ref(fig:simulations); column 1, row 2).
The endorsement model then predicts this is a rather good generic sentence (Figure\ \@ref(fig:simulations), bottom row) even though not all dogs bark (referent prevalence of barking among dogs shown with arrow in Figure\ \@ref(fig:simulations) row 2).

@@

A second observation regarding generics is that their truth conditions cannot be so lenient so to always convey *existential* quantification (e.g., "Some").
Indeed, it's intuitively plausible that some kangaroos do have spots but a rational language user might feel the generalization "Kangaroos have spots" is too strong to assert.
The model exhibits this same restraint, predicting a very low endorsement probability for this generic statement (Figure\ \@ref(fig:simulations); column 2).
The reason is that, given the interpretation model's background knowledge about the property, the statement "Kangaroos have spots" will be interpreted like "Dogs bark": it would mean *almost all* kangaroos have spots. 
The endorsement model, which believes that very few kangaroos have spots, then would rather not endorse the statement because it would mislead the listener.

**3. Sharks don't eat people.**

A similar logic can be applied to this example, which is the converse of the affirmative generalization "Sharks eat people", which is assumed true despite the low actual number of incidents involving sharks eating people [@Leslie2007]. 
The referent prevalence---the number of sharks that don't eat people---is thus very high: Almost all sharks don't eat people.
Yet, intuitively, the generalization is again misleading. 
Because for almost all animal categories, almost members of those categories don't eat people (i.e., the prevalence of the feature is almost 100% for all categories because very few things eat people), interpreting the statement "Sharks don't eat people" would lead one to believe that no sharks eat people, which is too strong.
It is interesting to note that because of the high referent-prevalence, the model is less certain of its decision, predicting an endorsement probability around 40%.
We will test this quantitative prediction in one of our experiments.

**4. and 5. Robins lay eggs vs. Robins are female**

@@

The prior distributions over the prevalence of both features are shown in Figure\ \@ref(fig:simulations) (columns 5 and 6).
As described above, the priors are different: Many animals have zero egg-layers (0% prevalence), while the vast majority of animal categories have female members in exactly the same proportion (50%).
Interestingly, the generic interpretation model returns roughly the same interpretation distribution for each hypothetical utterance: In each case, the model believes 50% have the property.^[
  This behavior might suggest a potential area for model revision. 
  Intuitively, "Xs are female" should mean "All Xs are female".
  In this paper, our empirical interest concerns endorsements or truth judgments; at this point, the interpretation model is only an intermediate representation useful for deriving predictions about endorsement.
]
However, only in the case of "Robins lay eggs" does the endorsement model actually assert the generic; it does so because the listener interpretation would more aligned with the referent prevalence than if the endorsement model produced the null utterance, which would instead return the prior.
In the case of "Robins lay eggs", the hypothetical generic utterance does not carry any information (the listener posterior is not different from the prior); in this case, however, the generic would not be misleading only uninformative, and hence the model predicts an endorsement probability of 0.5 (i.e., *neither true nor false*).

**6. Mosquitos carry malaria.**

Finally, the statement "Mosquitos carry malaria" is intuitively true in spite of the fact that the vast majority of mosquitos are actually malaria-free.
The uncertain-threshold model comes to the same conclusion---endorsing a generic when the referent prevalence is very low---because the listener interpretation distribution would be more aligned with the referent prevalence than the prior. 
Intuitively, this is because not that many animals have malaria-carriers among them and so in a sense, having malaria is diagnostic for being a mosquito, an intuition that is often arrived at with this example.
This example also highlights that the construct of *cue validity*---the probability of the feature given the category e.g., $P({is a mosquito} \mid \text{carries malaria})$---plays a role in our model. 
We return to this relationship in the next section, and provide a precise mathematical derivation of the relationship between the prevalence prior and cue validity in Appendix A.

@@

Imagine there is a job in the local bureaucrat's office to handle the mail from Antarctica and this job is assigned to Mary; to date, however, there has never been any mail from Antarctica [@Cohen1999]. 
That is, the statement "Mary *has handled* mail from Antarctica" is false.
Sentence (7) is still thought to be intuitively true despite zero actual instances of the event.
This highlights an important ambiguity in the theoretical commitments of the uncertain threshold model: The endorsement model aims to communicate some referent prevalence $p$, but does the prevalence represent the actual, objective frequency in the world or a subjective, predictive degree of belief in the head?

We posit that Sentence (7) is an extreme case of generalizations expressing *predictive* probabilities.
We expect this context sets up the expectation that any future mail coming from Antarctica will be handled by Mary. 
<!-- In Expt. 2c., we experimentally dissociated predictive probability from past frequency, finding that generalization endorsement tracked predictive probability, but not frequency. -->
<!-- We expect the same to be true with Mary: Any future mail coming from Antarctica will be handled by Mary.  -->
Thus the predictive probability that Mary will handle Antarctican mail, if there is any, is high.
Predictive probabilities often track past frequency or actual prevalence in the world, but people's internal models of how the world works can lead them to diverge. 
In this case our understanding of Mary's job leads to a strong predictive probability in the absence of past frequency evidence.

@@

Imagine that every current and past Supreme Court Justice has a social security number which is an *even number*. 
Sentence (8) is still considered false, even though the property holds for exactly 100\% of the category [@Cohen1999].
We predict the rejection of Sentence (8) is the result of abstract, intuitive theories guiding our subjective predictive probabilities.
That is, because observers strongly believe one's social security number does not influence selection for the Supreme Court, speakers would assign a roughly 50\% subjective probability that the *next* justice would have an even social security number.
Then, if we constructed the comparison class by considering other professionals, Sentence (8) would be rated by our model as similar to "Birds are female", because all professions have roughly the same probability of having employees with even social security numbers.

@@

\mht{both this and the "surprisng prediction" above are intutiions/examples we are not going to test empirically (however, here we at least give a model for it)... should these appear elsewhere, or somehow be marked upfront?}

Understanding how a semantic representation *composes* is another important test for a theory of generalizations in language.
@Nickel2008 suggests Sentence (9) is semantically equivalent to: "Elephants live in Africa and elephants live in Asia".
Statistical accounts of generics have trouble with this kind of conjunction: It cannot be the case that *most* (more than half) elephants live in Africa and *most* (more than half) elephants live in Asia, unless we are to posit *international elephants* (i.e., individual elephants who live part-time in Africa and part-time in Asia), which we would rather prefer not to posit.
However, this is the only interpretation available if the generalization is given a fixed, majority-based quantificational semantics, as posited in most statistical views of generics [e.g., if the generic means *most*; @Cohen1999].
Our theory does not provide a fixed semantics for generalizations, but an uncertain one, which can be updated as more information comes in.
In fact, with prior knowledge suggesting against the existence of international elephants and without any further semantic assumptions, our model interprets Sentence (9) as meaning some elephants live in Africa and that different ones live in Asia.^[
  An implementation of this example can be found: [http://forestdb.org/models/generics-conjunction.html](http://forestdb.org/models/generics-conjunction.html).
]
Interestingly, an incremental parsing of Sentence (9) (i.e., upon hearing only the first part of the sentence that "Elephants live in Africa") leads our model to believe that *most* elephants live in Africa.
When the sentence is completed ("...and Asia"), the model non-monotonically updates its beliefs to something weaker: *some* elephants live in Africa (and others in Asia, but few in other places).
The flexibility of our model accommodates new evidence that might otherwise contradict previous utterances because it maintains uncertainty about the precise meaning of the utterance.

@@

>>>>

Endorsement predictions depend upon the uncertain threshold model's interpretations, which are highly sensitive to background knowledge.
Background knowledge may be richly structured, reflecting intuitive theories about the underlying causes of different kinds of properties [cf., @Leslie2007].
This model posits, however, that the only impact of structured knowledge on truth judgments is in their implications for beliefs about prevalence, formalized in terms of a prevalence prior $P(p)$. 
\ndg{i’d save this for expt sections, and here just say that we posit intuitive priors for illustrations and introduce a method to actually elicit them below}
\mht{without this, we don't have an explanation for what the prevalence prior is, which is important for understanding the examples... i removed stuff about Betas, but check this again...}
The prevalence prior reflects a listener's beliefs about the prevalence of the feature in an unknown or unfamiliar category and be conceptualized as a distribution over the prevalence among alternative categories (Figure\ \@ref(fig:simulations)).
<!-- For example, think about five of your favorite kinds of animal: What percentage of each of them *are female*? What percentage of each of them *lay eggs*? -->
For example, the *lay eggs* prevalence prior should be bimodal, with substantial probability mass near 0 (since many animal categories do not have egg-layers) and a secondary component peaked around 0.5 (because among the animals with egg-layers, only the female members of the category have the property).
Conversely, the distribution over the prevalence of *being female* is unimodal and centered at 0.5, because almost all animals have female members in that proportion.
Figure\ \@ref(fig:simulations) (row 1) shows these and other intuitive prevalence priors for different properties (we introduce methods for empirically eliciting these priors in the experimental section).
We now turn our attention to the endorsement model's predictions.

@@

```{r workedExamples, results="asis"}
worked.examples <- data.frame(
  Example = c("1. Dogs bark",
              "2. Kangaroos have spots",
              "3. Robins lay eggs",
              "4. Robins are female",
              "5. Mosquitos carry malaria", 
              "6. Sharks don't eat people",
              "7. Mary handles the mail from Antarcita",
              "8. Supreme Court Justices have even social security numbers",
              "9. Elephants live in Africa and Asia"),
  `referent prevalence` = c("95%", "5%",  "50%", "50%", "5%", "85%", "0 times", "100%", "50%/50%"),
  `intuitive truth value` = c(
    "T", "F",  "T", "U",
    "T", "F","T", "F", "T" 
  ),
  `potential issue` = c(
    "Not all dogs bark",
    "Some kangaroos could have spots",
    "Only female robins (50%) lay eggs",
    "Same number that lay eggs (50%)",
    "Very few mosquitos actually carry malaria",
    "Most sharks don't eat people",
    "Consider there has never been any mail from Antarcita",
    "Consider 100% have even social security numbers",
    "Impossible for most to live in Africa and most to live in Asia"
  )
) %>%
  rename("Referent prevalence" = referent.prevalence,
         "Intuitive truth value" = intuitive.truth.value,
         "Issue" = potential.issue)

@@

Often the first observation with generics is that they appear to behave like universal quantifiers (e.g., "All") that permit exceptions: Not all dogs bark, but still "Dogs bark" is true.
The uncertain threshold account immediately accomodates this example because there is no single, fixed threshold beyond which a generic statement becomes true.
Rather, listeners have uncertainty about the threshold which leads to graded interpretations (e.g., "Dogs bark" probably means *almost all* dogs bark; Figure\ \@ref(fig:simulations); column 1, row 2).
The endorsement model predicts "Dogs bark" is a rather good generic sentence (Figure\ \@ref(fig:simulations), bottom row) even though there are exceptions to the universal generalization (referent prevalence of barking among dogs shown with arrow in Figure\ \@ref(fig:simulations) row 2).

@@

A second observation regarding generics is that their truth conditions cannot be so lenient so as to always convey *existential* quantification (e.g., "Some").
Indeed, it's intuitively plausible that some kangaroos do have spots but a rational language user might feel awkward to assert the generic "Kangaroos have spots".
The model exhibits this same restraint, predicting a very low endorsement probability for this generic sentence (Figure\ \@ref(fig:simulations); column 2).
The reason is that, given the interpretation model's background knowledge about the property, the statement "Kangaroos have spots" will be interpreted like "Dogs bark": it would mean *almost all* kangaroos have spots. 
The endorsement model, which believes that very few kangaroos have spots, then would rather not endorse the statement because it is too strong a statement and would mislead the listener.

**3. and 4. Robins lay eggs vs. Robins are female**

@@

The prior distributions over the prevalence of both features are shown in Figure\ \@ref(fig:simulations) (columns 5 and 6).
As described above, the priors are different: Many animals have zero egg-layers (0% prevalence), while the vast majority of animal categories have female members in exactly the same proportion (50%).
Given this background knowledge, the generic interpretation model returns roughly the same interpretation distribution for each hypothetical utterance: In each case, the model believes 50% have the property.
However, only in the case of "Robins lay eggs" does the endorsement model actually assert the generic; it does so because the listener interpretation would be more aligned with the referent prevalence than if the endorsement model produced the null utterance, which would instead return the prior.
In contrast, "Robins are female" does not carry any information (the listener posterior is not different from the prior); here, the generic would not be misleading but uninformative, and hence the model predicts an endorsement probability of 0.5 (i.e., *neither true nor false*).
Indeed, previous studies on generic endorsements have found "Robins are female" to receive an intermediate endorsement level [@Prasada2013].

**5. Mosquitos carry malaria.**

Finally, the statement "Mosquitos carry malaria" is intuitively true in spite of the fact that the vast majority of mosquitos are actually malaria-free.
The prevalence prior for "carries malaria" is highly skewed towards low prevalence levels: many animals do not have malaria-carriers among them and even for those that do, the prevalence is expected to be quite low.
Then, carrying malaria is significantly more true of mosquitos than other animals, an intuition that is often arrived at with this example.
This kind of behavior is related to the construct of *cue validity*---the probability of the feature given the category e.g., $P({is a mosquito} \mid \text{carries malaria})$---which we return to in the next section.
The uncertain-threshold model displays a critical behavior: It can endorse generics when the referent prevalence is very low.

<!-- and then point out that this is related to cue validity, as explored later. (or just wait to introduce cue validity until later.) -->
<!-- ---because the listener interpretation distribution would be more aligned with the referent prevalence than the prior.  -->
<!-- Intuitively, this is because many animals do not have malaria-carriers among them and so in a sense, having malaria is diagnostic for being a mosquito -->
<!-- We return to this relationship in the next section, and provide a precise mathematical derivation of the relationship between the prevalence prior and cue validity in Appendix A. -->

**6. Sharks don't eat people.**

When the prevalence of the feature is very high, it is also not necessary for the model to endorse the generic.
"Sharks don't eat people" is predicted to be a somewhat strange utterance, despite the fact that the vast majority of sharks do not eat people.^[
 We create this example as the converse of the classic example "Sharks eat people", predicted to be *true* despite low prevalence [@Leslie2007]. We reverse this example to show how the model deals with a high-prevalence feature that is predicted to be false or infelicitous.  
] 
<!-- Yet, intuitively, the generic would be misleading (similar in intuition to "Kangaroos have spots"). -->
Because for almost all animal categories, the prevalence of *not eating people* is almost 100% (very few things eat people), interpreting the statement "Sharks don't eat people" would lead one to believe that no sharks eat people, which is too strong.
It is interesting to note that because of the high referent prevalence, the model is less certain of its decision, predicting an endorsement probability around 40% (i.e., somewhat false).
We will test this quantitative prediction in Case Study 1 on generics.

@@

Imagine there is a job in the local bureaucrat's office to handle the mail from Antarctica and this job is assigned to Mary; to date, however, nobody has ever sent mail to the office from Antarctica [@Cohen1999]. 
In other words, the statement "Mary has handled mail from Antarctica" is false.
"Mary handles the mail from Antarctica" is still thought to be intuitively true despite zero actual instances of the event.
This highlights an important ambiguity in the theoretical commitments of the uncertain threshold model: The endorsement model aims to communicate some referent prevalence $p$, but does the prevalence represent the actual, objective frequency in the world (e.g., the number of times Mary has handled mail in the past) or a subjective, predictive degree of belief in the head (e.g., our prediction that were the appropriate situation to arise, Mary would be handling the mail from Antarctica)?

We posit that Mary the Antarctic mail handler is an extreme case of generalizations expressing *predictive* degrees of belief.
We expect this context sets up the expectation that any future mail coming from Antarctica will be handled by Mary. 
<!-- In Expt. 2c., we experimentally dissociated predictive probability from past frequency, finding that generalization endorsement tracked predictive probability, but not frequency. -->
<!-- We expect the same to be true with Mary: Any future mail coming from Antarctica will be handled by Mary.  -->
Thus the predictive probability that Mary will handle Antarctican mail, if there is any, is high.
Predictive probabilities often track past frequency or actual prevalence in the world, but people's internal models of how the world works can lead them to diverge. 
In this case our understanding of Mary's job leads to a strong predictive probability in the absence of past frequency evidence.
We explore this question experimentally in Case Study 2 on habituals.

@@

Imagine that every current and past Supreme Court Justice has had a social security number which was an *even number*. 
"Supreme Court Justices have even social security numbers" is still considered false, even though the property holds for exactly 100\% of the category [@Cohen1999].
We predict the rejection of even social security numbered Justices is the result of our intuitive theories guiding our subjective predictive probabilities, which feed into the endorsement model.
That is, we predict observers strongly believe there is no causal relation between the evenness of one's social security number and selection for the Supreme Court, and thus would assign a roughly 50\% subjective probability that the *next* justice would have an even social security number.
Then, "Supreme Court Justices have even social security numbers" would be rated by our model as similar to "Birds are female", because all professions have roughly the same probability of having employees with even social security numbers.
This example is thus the conceptual opposite of Mary the Antarctic mail handler, where our internal model of the world led us to a strong degree of belief in the property holding in the future; with the Supreme Court Justices, our internal models lead to to weak degree of belief in the property holding in the future.

@@

<!-- \mht{both this and the "surprisng prediction" above are intutiions/examples we are not going to test empirically (however, here we at least give a model for it)... should these appear elsewhere, or somehow be marked upfront?} -->

Understanding how a semantic representation *composes* is another important test for a theory of generalizations in language.
@Nickel2008 suggests that "Elephants live in Africa and Asia" is semantically equivalent to: "Elephants live in Africa and elephants live in Asia".
Statistical accounts of generics have trouble with this kind of conjunction because they historically have appealed to fixed, truth-conditional cutoffs for generics [e.g., generic means *more than half*; @Cohen1999].
It cannot be the case that most (more than half of) elephants live in Africa and *most* (more than half of) elephants live in Asia, unless we are to posit *international elephants* (i.e., individual elephants who live part-time in Africa and part-time in Asia), which are intuitively implausible.
<!-- However, this is the only interpretation available if the generalization is given a fixed, majority-based quantificational semantics, as posited in most statistical views of generics [e.g., if the generic means *most*; @Cohen1999]. -->
Our theory does not provide a fixed semantics for generics, but an uncertain one which can be updated as more information comes in.
In fact, with prior knowledge suggesting against the existence of international elephants, our model interprets "Elephants live in Africa and Asia" as meaning that some elephants live in Africa and that different ones live in Asia.^[
  An implementation of this example can be found: [http://forestdb.org/models/generics-conjunction.html](http://forestdb.org/models/generics-conjunction.html).
]
Interestingly, an incremental parsing of the sentence  (i.e., upon hearing only the first part of the sentence that "Elephants live in Africa") leads our model to believe that *most* elephants live in Africa.
When the sentence is completed ("...and Asia"), the model non-monotonically updates its beliefs to something weaker: *some* elephants live in Africa and others in Asia (but few in other places).
The flexibility of our model accommodates new evidence that might otherwise contradict previous utterances because it maintains uncertainty about the precise meaning of the utterance.
We leave for future work the testing of quantitative predictions for this kind of example.

@@

<!-- ^[ -->
<!--   This behavior might suggest a potential area for model revision.  -->
<!--   Intuitively, "Xs are female" should mean "All Xs are female". -->
<!--   In this paper, our empirical interest concerns endorsements or truth judgments; at this point, the interpretation model is only an intermediate representation useful for deriving predictions about endorsement. -->
<!-- ] -->
